Embedded Scalars, Preferred Readings \& Intonation: {A}n Experimental Revisit

### 2014-05-29 text/paper/paper4.tex
Conclusions sec com

\title{Embedded Scalars: {A}n Experimental Revisit} \author{Fabian
  Schlotterbeck, Michael Franke and Petra Augurzky} \date{}

@@

The data obtained by this method is problematic for both
traditionalism and grammaticalism. Contrary to traditionalism and the
intonational markedness hypothesis, we found evidence for local
readings for both kinds of critical sentences, irrespective of
intonational stress. \todo{enlarge: how big was this effect? can
  traditionalists explain this away?} But also grammaticalism is
incongruent with the observed pattern in our data. Backed up by the
Strongest Meaning Hypothesis
\citep{DalrympleKanazawa1998:Reciprocal-Expr}, grammaticalism predicts
local readings to be the preferred readings, but we found no evidence
for that. More strongly even, the preference order among putative
readings suggested by our data cannot even in principle be explained
by comparisons of logical strength among putative readings. In sum,
our data sheds doubt on both traditionalism and grammaticalism in
their currently most common forms. We conclude the paper by some
speculation about how either position might be amended in the light of
our data.

@@

The grammatical approach, as described so far, is not yet a
respectable theory. It makes little empirical commitments of
substance. It must be supplemented by a selection principle that
determines which of the many readings made available by the grammar
actually obtain or are preferred. Proponents of grammaticalism are
aware of this and have suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives. \todo{verify that this is the proposal} Another proposal
is that readings are preferred in relation to how well they answer the
contextually given question under discussion
\citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}. Yet
another proposal is to select readings by comparison of logical
strengths, as suggested by the \mymark{strongest meaning hypothesis}
of \citet{DalrympleKanazawa1998:Reciprocal-Expr}
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. We
will focus entirely on this latter selection principle, because it
seems to be the most prominent in the literature and because it makes
the clearest predictions. That means that all we will say about
grammaticalism in this paper strictly applies only to grammaticalism +
strength based selection of readings.

@@

A related objection would be that maybe it is not feasible to actually
compare results from the preference-related controls to results from
our target conditions, because the source of different readings is
fundamentally different. In the former we have a syntactic ambiguity,
in the latter a small number of competing putative pragmatic
enrichments. The former is a case of perceived ambiguity (some of our
subjects reported this), the latter arguably is not
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
if the distribution of responses is indicative of the known preference
patterns in preference-related controls, this would not necessarily
mean that the distribution of responses in target conditions is
indicative of preference relations.

Although we cannot rule out that the incremental verification task
detects preferences among multiple readings only for some sources of
variability in readings and not others, it is by no means clear why
this should be so. The burden of argument is on the sceptic. But even
if we assume that the incremental verification task is incapable of
reliably indicating preferences among candidate pragmatic enrichments
of the relevant kind, then it should at least be able to indicate
availability of readings. So if we deny the observed response patterns
the power to indicate preference, what should the information about
availability of readings be that we may safely infer from it? It
should be that local readings are available and that global readings
are not. But that is not predicted by either traditionalism or
grammaticalism. The former cannot explain the substantial number of
local responses, the latter doesn't explain the absence of global
responses.

\paragraph{Future Directions.} The data obtained from our experiment
also raises a number of issues for future research. Firstly, we found
a tendency of subjects to give literal or local answers rather
consistently. This seems to be related to what has been reported in a
number of previous studies on unembedded \emph{some}, where a tendency
was found as well in subjects to either consistently interpret
\emph{some} `semantically' as \emph{some and possibly all} or to
consistently interpret it `pragmatically' as \emph{some and not all}
\citep[e.g.][]{NoveckPosada2003:Characterizing-,BottNoveck2004:Some-Utterances,DegenTanenhaus2012:Processing-Scal}. This
inter-subjective consistency has not been paid any attention to in
previous studies on embedded scalar items. It would be interesting to
test whether there is any indication of a correlation between, on the
one hand, subjects' tendencies to give literal vs.~local answers for
embedded occurrences of \emph{some} and, on the other hand, subjects'
tendencies to interpret unembedded \emph{some} semantically or
pragmatically. If such a correlation exists, it would point towards
the possibility that different subjects may even have different
lexical representations of \emph{some} or possibly employ different
processes of interpreting it. If so, this would severely mix up the
theoretical landscape and should also be taken into account in any
further study on the processing of scalar items.

Secondly, given the scarce literature on the topic, more research into
the role of intonational markedness on pragmatic interpretation is
clearly needed. Claims about the correlation of certain
interpretations with certain patterns of intonation are frequently
made in the theoretical literature, but often only backed up by
introspection and not made precise in their commitment to the exact
kind of intonational pattern that is hypothesized to be relevant. The
intonational markedness hypothesis, for instance, seems \emph{prima
  facie} plausible, but we did not find any evidence for it in our
experiment. This much is not yet any direct evidence against it
either, because it is certainly possible that the contrastive accents
we realized in our experiments, although the most plausible and in
line with current theorizing about German contrast marking, was not
the right kind of intonational markedness. This shows for one that
claims about the putative effects of intonation and prosody must be
more carefully phrased and filled with detail, but also points to
further empirical research. The combination of production experiments
the output of which is fed into a subsequent interpretation experiment
comes to mind as a first natural continuation.

@@

We tested experimentally the predictions of traditionalism and
grammaticalism with respect to the availability of and preferences
over three types of conceivable pragmatic readings of two types of
sentences, in which scalar \emph{some} occurred in the scope of either
a monotonic quantifier \emph{all} or the non-monotonic quantifier
\emph{exactly one}. To avoid potential confounds of the pictorial
material, such as through typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi},
we employed an incremental verification task \citep{Conroy2008}, in
which subjects were presented with partially covered pictures and
asked to uncover the picture sequentially until they felt able to give
a categorical truth-value judgements. Sentence material was presented
auditorily and contrastive stress on embedded \emph{some} was
manipulated so as to test the intonational markedness hypothesis,
which is often used to supplement traditionalism. Both traditionalism
and grammaticalism do not predict our data well.  Traditionalism fails
to account for the non-negligible number of responses indicating local
readings even for intonationally unmarked sentences. Grammaticalism
fails to predict the absence of answers indicative of global readings
and the relative abundance of answers indicative of literal readings.

>>>>

\title{Embedded Scalars, Preferred Readings \& Intonation: {A}n
  Experimental Revisit} 
\author{% Fabian
  % Schlotterbeck, Michael Franke and Petra Augurzky
} 
\date{}

@@

The data obtained by this method is problematic for both
traditionalism and grammaticalism. Contrary to the predictions of
traditionalism, we observed many answers indicative of local readings
of \es-sentences, irrespective of intonational markedness. But also
grammaticalism is incongruent with the observed pattern in our
data. Backed up by the Strongest Meaning Hypothesis
\citep{DalrympleKanazawa1998:Reciprocal-Expr}, grammaticalism predicts
local readings to be the preferred readings, but we found no evidence
for that. More strongly even, the preference order among putative
readings suggested by our data cannot even in principle be explained
by comparisons of logical strength among putative readings. In sum,
our data sheds doubt on both traditionalism and grammaticalism in
their currently most common forms. We conclude the paper by some
speculation about how either position might be amended in the light of
our data.

@@

The grammatical approach, as described so far, is not yet a
respectable theory. It makes little empirical commitments of
substance. It must be supplemented by a selection principle that
determines which of the many readings made available by the grammar
actually obtain or are preferred. Proponents of grammaticalism are
aware of this and have suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives. Another proposal is that readings are preferred in
relation to how well they answer the contextually given question under
discussion
\citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}. Yet
another proposal is to select readings by comparison of logical
strengths, as suggested by the \mymark{strongest meaning hypothesis}
of \citet{DalrympleKanazawa1998:Reciprocal-Expr}
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. We
will focus entirely on this latter selection principle, because it
seems to be the most prominent in the literature and because it makes
the clearest predictions. That means that all we will say about
grammaticalism in this paper strictly applies only to grammaticalism +
strength based selection of readings.

@@

A related objection is that it is not feasible to actually compare
results from the preference-related controls to results from our
target conditions, because the source of different readings is
fundamentally different. In the former we have a syntactic ambiguity,
in the latter a set of putative pragmatic enrichments. The former is a
case of perceived ambiguity (some of our subjects reported this), the
latter arguably is not
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
if the distribution of responses is indicative of the known preference
patterns in preference-related controls, this would not necessarily
mean that the distribution of responses in target conditions is
indicative of preference relations. Although we cannot rule out that
the incremental verification task detects preferences among multiple
readings only for some sources of variability in readings and not
others, it is by no means clear why this should be so. The burden of
argument is on the sceptic. But even if we assume that the incremental
verification task is incapable of reliably indicating preferences
among candidate pragmatic enrichments of the relevant kind, then it
should at least be able to indicate availability of readings. So if we
deny the observed response patterns the power to indicate preference,
what should the information about availability of readings be that we
may safely infer from it? It should be that local readings are
available in \es-sentences and that global readings are unattested in
\as- and \es-sentences. But that alone is not predicted by either
traditionalism or grammaticalism. The former cannot explain the
substantial number of local responses in \es-sentences, the latter
does not explain the absence of total global responses.

Another line of criticism that challenges the validity of
(\ref{bsp:observed-preferences}) targets the Bayesian
salience-competition model that we used to derive it. We readily admit
that our model is highly simplistic. Several aspects are especially
noteworthy here. Firstly, we did not distinguish different types of
errors, and did not take into account that when errors occurred
earlier in the sequence, later decision points could no longer be
reached. This way the model may actually have unduly de-emphasized the
salience of readings that can be judged later in a sequence. Another,
more complicated model of the likelihood of answer patterns, might
therefore have given support to different conclusions. Unfortunately,
it is not clear how exactly such a more complex generating model
should be set-up. Eventually, a full processing model, design for data
from incremental verification tasks would be needed. We cannot offer
such a model here, but can only the offer the model that we did use as
a first step in that direction. More importantly, though, the only
part of our conclusions that might plausibly change in
(\ref{bsp:observed-preferences}) for a different generating model is
that the 3 \emph{false} answers in the \as-conditions with neutral
intonation might support the conclusion that the local reading is
marginally attested. But this hypothetical effect will certainly be no
bigger than marginal, and so no genuinely different conclusions of
theoretical relevance should be expected.

@@

We tested experimentally the predictions of traditionalism and
grammaticalism with respect to the availability of and preferences
over three types of conceivable pragmatic enrichments of two types of
sentences, in which scalar \emph{some} occurred in the scope of either
a monotonic quantifier \emph{all} or the non-monotonic quantifier
\emph{exactly one}. To avoid potential confounds of the pictorial
material, such as through typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi},
we employed an incremental verification task \citep{Conroy2008}, in
which subjects were presented with partially covered pictures and
asked to uncover the picture sequentially until they felt able to give
a categorical truth-value judgements. In order to control for
potential confounds of ``silent prosody,'' sentence material was
presented auditorily and contrastive stress on embedded \emph{some}
was manipulated. This way we were able to also explicitly test the
intonational markedness hypothesis, which is often used to supplement
traditionalism. Additionally, we included preference-related control
conditions in order to be able to deduce latent preferences for
candidate readings from the answer patterns we gathered. To this end,
we introduced a generative Bayesian model that gives probabilistic
predictions about possibly observable answer patterns in our task,
given concrete instances of relatives saliences of readings and other
parameters. Using the observed data, in particular in connection with
the preference-related control conditions, we inferred a posteriori
likely values of the latent cognitive parameters.

Backed up by our analyses, we concluded that both traditionalism and
grammaticalism do not predict our data well. Traditionalism fails to
account for high numbers of responses indicating local readings even
for intonationally unmarked \es-sentences. Grammaticalism, insofar as
it builds on disambiguation by logical strength, fails to predict the
absence of answers indicative of global readings and the relative
abundance of answers indicative of literal readings.

The clearest upshot of theoretical importance of this study is that
strength-based meaning selection does not seem to be a good
disambiguation criterion for grammaticalism. We tentatively suggested
an alternative selection principle for grammaticalism in terms of
``parsing economy.'' Another important conclusion from our data is
that \es-sentences, using non-monotonic quantifier \emph{exactly one},
appeared to be prone to receive a ``shallow interpretation.'' This may
be used by traditionalism to explain answers indicative of local
readings for \es-sentences. But the point is relevant in general. The
possibility of ``shallow interpretation'' of non-monotonic quantifiers
should be considered also when interpreting the results of other
experimental designs \citep[e.g.][]{CliftonDube2010:Embedded-Implic,
  ChemlaSpector2010:Experimental-Ev}.

On the methodologically side, we have argued that auditory
presentation of sentence material should be mandatory if prosodic
information is hypothesized to influence availability and preferences
of readings. We also argued that theoretical positions must be
formulated in such a way that they make testable predictions about
reading preferences where different potential readings are allowed
(contra a grammaticalism without commitment to any testable
disambiguation principle). We should then try to obtain information
about actual preference patterns by explicit comparison with
structures whose preference structures are well-studied, ideally with
some model of the data generating process for optimal data analysis.
